---
title: "Minimal NIfTI MRS reader example for R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this document is to demonstrate how [NIfTI MRS](https://docs.google.com/document/d/1tC4ugzGUPLoqHRGrWvOcGCuCh_Dogx_uu0cxKub0EsM/edit?usp=sharing) may be imported into the [`R`](https://www.r-project.org/) programming language, and perform basic processing steps with just few lines of code. Whilst dedicated MRS analysis packages, such as [`spant`](https://martin3141.github.io/spant/), are recommended for typical MRS analysis tasks in [`R`](https://www.r-project.org/), this example code is provided to aid software developers interested in writing their own tools to process NIfTI MRS files.

Firstly we download an example file, available from the [NIfTI MRS github repository](https://github.com/wexeee/mrs_nifti_standard), and load the excellent [`RNifti`](https://github.com/jonclayden/RNifti) library to read the data:

```{r read_nifti}
fname <- "csi.nii.gz"
if (!file.exists(fname)) {
  download.file("https://dl.dropboxusercontent.com/s/g5jwikq7btikdtt/csi.nii.gz?dl=0", mode = "wb", fname)
}
library(RNifti)
nifti_mrs <- readNifti(fname)
```

We might be interested to know the FID sampling frequency,

```{r get_fs}
1 / pixdim(nifti_mrs)[4]
```

and spatial transform to map our MRS data onto other MR images:

```{r get_xform}
xform(nifti_mrs)
```

a full header listing may also be output for inspection:
```{r get_header}
niftiHeader(nifti_mrs)
```

MRS data processing requires information beyond what is defined in the standard header, therefore we make use of the NIfTI MRS header extension (ecode 44), where additional parameters are stored in JSON format. Here we load the [`jsonlite`](https://arxiv.org/abs/1403.2805) library and output the first few items:

```{r read_ext}
ext_str  <- extension(nifti_mrs, 44, "character")
library(jsonlite)
ext_json <- fromJSON(ext_str)
head(ext_json)
```

To plot a spectrum we simply treat `nifti_mrs` as a multidimensional array, with the first three dimensions being the x, y and z voxel coordinates (8, 8, 12) and the fourth dimension as the FID. Once the FID has been extracted, we perform a Fourier transform and shift the data with [`pracma::fftshift`](https://cran.r-project.org/web/packages/pracma/index.html) to show 0 Hz in the centre of the spectrum:

```{r}
library(pracma)
fid  <- nifti_mrs[8, 8, 12, ]
spec <- fftshift(fft(fid))
plot(Re(spec), type = "l", xlab = "frequency [a.u.]", ylab = "intensity [a.u.]")
```

Finally, basic phasing and line-broadening steps may be performed to aid visual interpretation:

```{r}
fid_proc  <- fid * exp(-2i - 5e-3 * seq(0, dim(nifti_mrs)[4] - 1))
spec_proc <- fftshift(fft(fid_proc))
plot(Re(spec_proc), type = "l", xlab = "frequency [a.u.]", ylab = "intensity [a.u.]")
```